Decarbonisation options in the industry sector and conceptual model of industry in an integrated energy system model

### 2018-01-25 article.tex
status thursday on the way to the train

\usepackage{epstopdf}
\usepackage{lineno,hyperref}
\usepackage{booktabs}
\usepackage{float}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp}
\modulolinenumbers[5]

@@

\title{Fuel switch, savings, electrification - Decarbonisation in industry OR Pathways to Carbon Neutral Industrial Sectors: Integrated Modelling Approach with High Level of Detail for End-use Processes OR Decarbonization options in industry and conceptual model of industry in the integrated energy system}

@@

\section{Introduction}
% State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.
In light of the recent Paris Agreement, the necessity to aim at carbon neutral societies and switch to 100\% renewable energy systems is a clear goal. 
As many of the past international protocols, the summit in Paris highlighted the importance of decarbonise energy intensive sectors, reducing green house gasses (GHG) emissions. 
In Europe, similarly to other countries, two sectors stands out for substantial energy consumption and GHG emissions: power and industry sector. 
The power sector is historically associated with great use of energy resources because of the nature of most of the energy conversion processes (e.g combustion of fossil fuels). In the last years, the energy sector has been subject to changes in the ways to produce energy, with more technologies based on renewable energy resources gradually replacing fossil fuels-based plants. The increasing share of clean technologies in the power mix, along with other interventions (e.g. energy efficiency or sustainable fuels), has been (and it's still) driving down the level of GHG emission associated with the power sector.

Likewise, the industry sector is going through a similar path. 
With more than 125 TWh of electricity consumption, 851 TWh of fossil fuels used for energetic purposes and 671 TWh of fossil fuels feedstock, in 2010 industry sector used almost 30\% of the total EU energy consumption; the related GHGs corresponded to 9\% of the total EU28 emissions stock\cite{Eurostat2017,Lechtenbohmer2016}.
As for the power sector, technical options are available to decarbonise the industry. The three main established categories are (i) electrification, (ii) fuel substitution and (iii) improved energy efficiency \cite{Lechtenbohmer2016, Akashi2011, IRENA2014}.

%ELECTRIFICATION
Given the increasing share of renewable energy technologies in the energy systems, electrified processes powered by renewable electricity are expected to replace both fossil fuels used as energy and as feedstock in a near future \cite{Lechtenbohmer2015, EnergyCommission2017, Energikommissionen2017, Lin2017}. \cite{Lechtenbohmer2016} have estimated that, in a simulated 2050 scenario, a deep decarbonisation via electrification of the industry can lead up to almost 300 Mt of $CO_2$ direct savings for the EU28. Also, \cite{IRENA2014} projects a future increase of the share of electricity use in the industrial sector from about 10\%-15\% in 1980s, to 25\%-30\% by 2030.
\\
%FUEL SUBSTITUTION
Fuel substitution is also widely considered for future scenarios \cite{Lin2016,Rehfeldt2016}. \cite{Iea2012} considers that switching fuels and feedstock to renewable fuels (e.g. biogas or hydrogen) can account for 300 $MtCO_2$-eq of emission reduction in a 4 degrees scenario. Similar analyses performed in Italy \cite{Enea2016} and England \cite{Pye2015}, report fuel switching as a main pillars for the transition to a decarbonised energy future.
\\
%ENERGY EFFICIENCY
Furthermore, although studies have acknowledge that the industrial energy efficiency is generally increasing for most of the processes, there is still a large unexploited potential, particularly for industries characterised by high-temperature processes. On a Danish context, \cite{Buhler2016} shows that food, paper and chemical industry are more efficient compared to others processes, characterised by high temperatures. Policies in place in different countries, e.g. China \cite{Li2017,Zheng2017,Lin2017a,Miao2018,Guo2017}, Italy\cite{MazziottiGomezdeTeran2017}, Austria \cite{Karner2015} or Slovenia \cite{Pusnik2017} have proven to be an effective tool to reduce emission from industrial energy-related use, fostering both technology innovation and emission reduction.  
\\

In this context of international industry transformation, Denmark is striving to find solutions to reach the 2020 goals \cite{EnergyCommission2017,Energikommissionen2017}. 
On the Danish case, examples showed that use of heat waste \cite{Buhler2017a}, the electrification \cite{DanishEnergyAgency2014a} and the adoption of energy efficiency interventions in the industry sector \cite{DEAeff18,Buhler2016} can lower the use of fossil fuels and the related GHG emissions. Furthermore, considering the recent developments of generation in Denmark, renewable gases have the potential to replace natural gas to some extent \cite{Schweitzer2017,Lise2017,Jensen2017}.
Although the response of these singular studies has been promising, none of the investigations considered a proper system-wide impact resulting from changes in the industry sector. 
As both the power and industry, the most intensive energy consuming sectors, are highly interrelated, applying decarbonisation measures in one sector can result in an impact on the other. 
An energy system view/analysis, employing suitable tools like TIMES \cite{IEA-ETSAP}, EnergyPlan \cite{DepartmentofDevelopmentandPlanning}, Balmorel \cite{balmorel} or Prime \cite{NationalTechnicalUniversityofAthens}, is thus necessary to assess the size and the scale of such impact.

Nonetheless, although industry covers a relevant share of energy usage, the industrial sector is rather under-represented in most of the energy system models, particularly when investigating scenarios on decarbonisation pathways.

For the sake of exploring industrial decarbonisation options enforcement and the related transformation of the Danish energy system, a detailed model of the industry sector is necessary. Details should thus consider not only the structure of the processes with input fuels, temperature heat levels and fuel usage characteristics, but also temporal profiles and geographical location of energy consumption.

The goal of the paper is thus threefold: 1) provide a representation of the Danish industry characteristics highlighting the potential for decarbonisation options, 2) propose a method to model the industrial sector, with fairly high level of details, within an energy system model and 3) present preliminary results of applying decarbonisation options in the Danish energy system.

@@

The remainder of the paper is structured as follows: \autoref{inddescr} provides an insight on the Danish industry sector, \autoref{meths} presents the methodologies adopted and \autoref{datadescr} describes the dataset used for the study. \autoref{outcdisc} reports the empirical results and provides a discussion on the outcomes and \autoref{endofpaper} concludes the paper highlighting the relevant findings and suggesting future developments on the study.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In the literature, industry is the term that commonly refers to a diverse variety of processes characterised by high energy intensity and use. As the term does not have a standard definition, the processes considered to represent the industrial sector vary from study to study, making similar analyses on the field quite difficult to compare.
\cite{Buhler2016} shows that, compared to other studies which considered different numbers of processes and industries accumulated in the industry sector \cite{Oladiran2007,Al-Ghandoor2010,Dincer2003,Sanaei2012}, the results derived from a more thorough approach provide additional insights and a stronger reliability. In this study, we will thus consider that industry includes agriculture, manufacturing and services; however we do not consider transportation, both as end-use and sector, within the industry sector.

The Danish national version of EU's nomenclature (NACE) is the Dansk Branchekode DB07 (Danish Branch Code), a statistical classification of economic activities that categorise each enterprise based on its main activity \cite{DanmarksStatistik2013}. 
For the groups of economic activities, detailed data on yearly energy usage by fuel type is available. A detailed clustering on the end-use of the fuels utilised is provided in \cite{VM2015}, but just for 57 groups excluding public service, mining and construction. The allocation of economic activities to the sectors mainly follows the grouping from Statistics Denmark; the exact allocation can be found in the Annex of \cite{VM2015}. As refinery, public service and construction are not included in the list of sectors for which detailed end-use data is provided \cite{VM2015}, they will not be included in this study.
\\
The consumption of the 57 sector is available as a total of 22 end-use processes supplied by 20 types of fuels. The distribution of fuels amongst the process categories is also provided \cite{VM2015}.
\\
The 22 energy end-uses can be clustered in four groups: \textit{internal conversion losses}, \textit{process heating} (heating/boiling, drying, inspissation, distillation, burning/sintering, melting/casting, other process heating up to/over 150 C), \textit{space heating} and \textit{utilities} (energy for heat pumps, lightning, pumping, comfort cooling, cooling/freezing, comfort ventilation, blowers, compressed air, hydraulic machinery, other electric motors, IT and electronics, other electric consumption).
\\
20 fuel types supplying processes include coal, fuel oil, petroleum coke, coke, diesel, gasoline (unleaded and coloured), LPG, natural gas, gas-oil, bio-gas, biomass (wood chips, wood pellets and straw), bio-oil, waste, electricity, district heating, solar heating and heat pumps.
\\
To provide a simpler visualisation of data, the 57 industrial sectors are clustered in five groups according to similarities in temporal patterns of energy consumption: agriculture, production single shift, production double shift, production triple shift and services. The underlying assumptions behind the classification are based on \cite{VM2016,Wiese2017}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy use in the Danish Industry}

@@

COMMENTS FOR THE FIGURE:
\begin{itemize}
    \item Remove transport from fuels (transport and transport machinery) CHECK
    \item Maybe report the percentages?
    \item One image with bars and legend
\end{itemize}

It´s relevant to notice that the fuel consumption among the sectors is distributed un-evenly, with electricity (36\%), natural gas (21\%), district heating (15\%), gasoil/kerosene (6.4\%) and petroleum coke (4\%) covering the majority of the consumption, while the other fuels contribute partially.

\iffalse
\autoref{heatmapagri}, \autoref{heatmapprod}, \autoref{heatmapserv} provide an indication on the fuel-use for the different industrial processes. The production data are reported as a single aggregated group; for the breakdown, please refer to (APPENDIX).

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_agri.png}
\caption{Breakdown of energy consumption by end-use and fuel in agriculture\cite{VM2015}}
\label{heatmapagri} 
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_prod.png}
\caption{Breakdown of energy consumption by end-use and fuel in production\cite{VM2015}}
\label{heatmapprod} 
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_serv.png}
\caption{Breakdown of energy consumption by end-use and fuel in services\cite{VM2015}}
\label{heatmapserv} 
\end{figure}
\fi

The heat maps, presented in three clustered groups, provide a useful indication on the most energy-intensive processes and on the most employed fuels. 
In agriculture, most of the fuel consumption is related to heating up to 150\textdegree (51\%); in production, drying (15\%), heating/boiling (16\%) and space heating (11\%) are predominant; while in service, space heating (50\%) and lightning (17\%) together consumer more than half of the total fuels used.

%top up with data
In production, there is a high diversity of end-use while the role of gas as a fuel is predominant. 
Space heating and transport dominate the total energy use in the service sector. While most space heat already supplied by district heat, natural gas is the second.

Furthermore, the data show an un-even distribution of fuel consumption in the various groups, with electricity (34\%) and gasoil/kerosene (25\%) being predominant in agriculture, electricity (31\%) and natural gas (31\%) in production, and electricity (50\%) and district heating (37\%) dominating the service sector.

This preliminary description of the Danish industry poses the bases for the modelling of the sector and the investigation of decarbonisation options. Indeed, with a clear overview of the end-use processes and the related fuel consumption, the industry sector can be conveniently integrated in an energy system model. 
The following section proposes a mathematical model for the industry sector and investigates on extensions to analyse decarbonisation options.

@@

\section{Methodology} \label{meths}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Balmorel, an energy system model}

The use of energy system models is useful to assess the role of industry in achieving pathways to decarbonised energy systems. As commodities like electricity and heat are highly related, in both power and industry sector (e.g. due to the use of heat pumps), only a simultaneous consideration can reflect the cost-minimal solutions from a system perspective. This allows to investigate what is usually defined as "socio-economic welfare maximisation" under different conditions. 
Among the tools available in the literature \cite{Connolly2010}, the energy system model Balmorel is adopted for the analysis \cite{balmorel}.
Balmorel is an energy system, linear programming, open source model that optimises investments and operation of power plants, storage devices and transmission lines in an international perspective \cite{Ravn2001}. 
The model considers a set of neighbouring countries operating in an interconnected electricity market. Each country is composed by one or several regions. Electricity can be traded and transmitted between adjacent regions, with limits imposed by given transmission capacity. 
To consider the heat sector, each electricity region is divided into several district heating areas. Heat transmission among such areas is not allowed.  
\\
Balmorel considers time according to years, seasons (weeks) and individual time units (hours). 
The user can freely decide the time horizon of the analysis and the choice depends on the needs for the specific investigation.
\\
The model allows to simulate scenarios where demand and supply of electricity and heat are balanced with operations and investments, considering local generation vs. import/export, price elasticity of the demand and other characteristics typical of the energy systems \cite{Munster2012}. 
\\
The model relies on a set of exogenous input data, including capacities of electricity and heat generation technologies, transmissions lines and heat and power demand. 
Energy generating technologies include CHP (both back-pressure and extraction), heat pumps, storage devices (for electricity and heat), and renewable based production technologies (hydro, wind, solar).
Additional key assumptions relate to fuel prices, CO2 costs, taxes and support schemes are required for specific analysis.
The original data available in the model can be modified for personal application in specific analyses. This allows Balmorel to be adapted for specific investigations that are not presently covered by the standard version of the model \cite{Wiese}.
\\
The model has been previously applied for a wide range of studies, such as integration of renewable technologies in the energy mix, analysis of market conditions, policies implementation, future role of district heating and impact of energy efficiency technologies in the energy system \cite{Ball2007,Jensen2008,Karlsson2008,Munster2012,Munster2010,Baldini2016a}.
In the next section, we will propose a modelling representation of the industry sector, as the Balmorel model lacks a comprehensive representation of the industry. Enhancing the model with details about industrial energy use and processes will improve the quality of the analyses and provide more details on the impact, in the energy system, of changes in the industrial sector.

@@

The energy system model Balmorel considers the industry sector only as a part of the total heat and electricity consumption within the optimisation. Thus all the end use processes and fuel specification, previously described, are currently neglected. The black dotted line in \autoref{ensyst} represents the boundaries in the current version of Balmorel, where the industry is considered simply as part of the total energy demand .

@@

\autoref{ensyst} shows the schematic of the novel structure proposed, to consider the diverse and rich characteristics of the industry sector in the energy system model. The same structure is applied for each of the 57 sectors considered.

The first addition introduced concerns the commodities part of the energy demand: electricity, fuels and heat.
The electricity demand now satisfies, in every single sector, the demand of the single end-uses. This addition facilitates the investigation of decarbonisation options (e.g. savings) at a later stage, as the new options can be applied directly to each end use (\textbf{maybe rewrite, the point here is that the more disaggregated is the consumption, the easier will be to investigate other options to satisfy the demand of decrease the demand}). 
Furthermore, electricity can be either be imported by the grid, or it can be self-produced using small plants. It can also be used to run heat pumps and satisfy the heat demand, or "elevate" the quality of heat, for peculiar processes. 
Heat, previously considered only as district heat demand, now includes three temperature levels: space heating, process heat low temperature and process heat high temperature processes. As an example how processes are categorised to these heat groups, Table \ref{tab:temperature_levels} summarises the categorisation is done in the case study.

Electricity, heat and fuels are considered as energy inputs for the end-use processes in industry. 
Fuels can also be used as fuel-feedstock directly for the industrial processes in the model to better represent reality. 
Heat is separated as space and process heat and distinguished between different temperature levels. In this way, it can be supplied by different sources and technologies.
Specific electricity profiles for the industrial sectors are the basis for the demand profiles fed into the model. Space heating for the industry sector can be supplied by district heating, but can also be self-generated, using the fuel in a plant on site to generate electricity and heat.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modelling of decarbonisation options}

\paragraph{Electrification}

\paragraph{Fuel substitution}

\paragraph{Excess heat}

\paragraph{Energy savings}

%Provide sufficient details to allow the work to be reproduced by an independent researcher. Methods that are already published should be summarized, and indicated by a reference. If quoting directly from a previously published method, use quotation marks and also cite the source. Any modifications to existing methods should also be described.
Industry plays a decisive part in the integrated energy system. The following energy system analysis questions appear:
\begin{itemize}
    \item Saving potential
    \item Fuel shift potential to gas, renewable gas
    \item Electrification potential
    \item Flexibility potential
    \item Usage of excess heat
    \item Usage of district heating
\end{itemize}
For an integrated modelling approach, a high level of detail required for
\begin{itemize}
    \item Fuel Usage
    \item End-use: electricity, heat specified by temperature levels
    \item Time: Temporal demand profiles
    \item Geography: Vicinity to heat demand / district heat network
    \item Industry group
\end{itemize} 
% \begin{itemize}
%     \item Overview whole model (Balmorel)
%     \item Zoom into industry add on (Figure)
%     \item Describe concept behind industry add on and heat transfer including formulas
%     \item Level of detail: Geographically, technologies, end-use
%     \item Data derivation (Could also be an own section or subsection)
% \end{itemize}

>>>>

\usepackage{epstopdf}
\usepackage{lineno,hyperref}
\usepackage{booktabs}
\usepackage{float}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp}
\usepackage{eurosym}
\modulolinenumbers[5]

@@

\title{Decarbonisation options in the industry sector and conceptual model of industry in an integrated energy system model}

@@

\section{Introduction}

In light of the recent Paris Agreement, the necessity to aim at carbon neutral societies and switch to 100\% renewable energy systems is a clear goal. 
As many of the past international protocols, the summit in Paris highlighted the importance to decarbonise energy intensive sectors, reducing greenhouse gas (GHG) emissions. 
In Europe, similar to other countries, two sectors stand out for substantial energy consumption and GHG emissions: power and industry. 
The power sector is historically associated with great use of energy resources because of the nature of most of the energy conversion processes (e.g combustion of fossil fuels). In the last years, the energy sector has been subject to changes with regard to the sources to produce energy. More technologies, based on renewable energy resources, are gradually replacing fossil fuels-based plants. The increasing share of clean technologies in the power mix, along with other interventions (e.g. energy efficiency or sustainable fuels), has been (and is still) reducing the level of GHG emission associated with the power sector.

Likewise, the industry sector starts and is required to follow a similar pathway. 
With more than 125 TWh of electricity consumption, 851 TWh of fossil fuels used for energetic purposes and 671 TWh of fossil fuels feedstock, the industry sector accounted for almost 30\% of the total EU energy consumption in 2010; the related GHGs corresponded to 9\% of the total EU28 emissions stock \cite{Eurostat2017,Lechtenbohmer2016}.
As for the power sector, technical options are available to decarbonise industry. The three main established categories are (i) electrification, (ii) fuel substitution and (iii) improved energy efficiency \cite{Lechtenbohmer2016, Akashi2011, IRENA2014}.

%ELECTRIFICATION
Given the increasing share of renewable energy technologies in the energy systems, electrified processes powered by renewable electricity are expected to replace both fossil fuels used as energy and as feedstock in the near future \cite{Lechtenbohmer2015, EnergyCommission2017, Energikommissionen2017, Lin2017}. \cite{Lechtenbohmer2016} have estimated that, in a simulated 2050 scenario, a deep decarbonisation via electrification of the industry can lead up to almost 300 Mt of $CO_2$ direct savings for the EU28. Also, \cite{IRENA2014} projects a future increase of the share of electricity use in the industrial sector from about 10\%-15\% in 1980s, to 25\%-30\% by 2030 of the total energy consumption.

%FUEL SUBSTITUTION
Fuel substitution is also widely considered for future scenarios \cite{Lin2016,Rehfeldt2016}. \cite{Iea2012} considers that switching fuels and feedstock to renewable fuels (e.g. biogas or hydrogen) can account for 300 $MtCO_2$-eq of emission reduction even in a 4 degrees scenario. This number would increase for more ambitious required scenarios. Similar analyses performed in Italy \cite{Enea2016} and England \cite{Pye2015}, report fuel switch as a main pillar for the transition to a decarbonised energy future.

%ENERGY EFFICIENCY
Furthermore, although studies have acknowledged that the industrial energy efficiency is generally increasing for most of the processes, there is still a large unexploited potential, particularly for industries characterised by high-temperature processes. In a Danish context, \cite{Buhler2016} shows that food, paper and chemical industry are more efficient compared to others processes, characterised by high temperatures. Energy efficiency policies in place in different countries, e.g. China \cite{Li2017,Zheng2017,Lin2017a,Miao2018,Guo2017}, Italy \cite{MazziottiGomezdeTeran2017}, Austria \cite{Karner2015} or Slovenia \cite{Pusnik2017} have proven to be an effective tool to reduce emission from industrial energy-related use, fostering both technology innovation and emission reduction.

In this context of international industry transformation, Denmark is striving to find solutions to reach the 2020 goals \cite{EnergyCommission2017,Energikommissionen2017}. 
With regard to Denmark, examples show that use of heat waste \cite{Buhler2017a}, electrification \cite{DanishEnergyAgency2014a} and adoption of energy efficiency interventions in the industry sector \cite{DEAeff18,Buhler2016} can lower the use of fossil fuels and related GHG emissions. Furthermore, considering recent developments of generation in Denmark, renewable gases have the potential to replace natural gas to some extent \cite{Lise2017,Jensen2017}.
Although these studies indicate promising possibilities, none of the investigations considered the system-wide context of changes of energy supply in the industry sector. 
As both the power and the industry sector are highly interrelated, applying decarbonisation measures in one sector can result in an impact on the other. 
An energy system view/analysis, employing suitable tools like TIMES \cite{IEA-ETSAP}, EnergyPlan \cite{DepartmentofDevelopmentandPlanning} or Balmorel \cite{Wiese2018, balmorel} is thus necessary to assess the size and the scale of these interdependencies.

Although industry covers a relevant share of energy usage, the industrial sector is rather under-represented in most energy system models, particularly when investigating scenarios on decarbonisation pathways.

For the exploration of industrial decarbonisation options enforcement and the related transformation of the Danish energy system, a detailed model of the industry sector is required. It should not only consider the structure of the processes with regard to input fuels, temperature heat levels and fuel use characteristics, but also temporal profiles and geographical location of energy consumption.

The goal of the paper is thus threefold: 1) provide a representation of the Danish industry characteristics highlighting the potential for decarbonisation options, 2) propose a method to model the industrial sector, with fairly high level of detail, within an energy system model and 3) present preliminary results of applying decarbonisation options in the Danish energy system.

@@

The remainder of the paper is structured as follows: \autoref{inddescr} provides an insight on the Danish industry sector, \autoref{meths} presents the methods adopted and \autoref{datadescr} describes the input data and its processing used for the study. \autoref{outcdisc} reports the empirical results and provides a discussion on the outcomes and \autoref{endofpaper} concludes the paper highlighting the relevant findings and suggests future research.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In the literature, industry is the term that commonly refers to a diverse variety of processes characterised by high energy intensity and use. As the term does not have a standard definition, the processes considered to represent the industrial sector vary from study to study, making similar analyses on the field difficult to compare. \cite{Buhler2016} shows that, compared to other studies which considered a restricted number of processes and industries \cite{Oladiran2007,Al-Ghandoor2010,Dincer2003,Sanaei2012}, the results derived from a broader approach provide additional insights and a stronger reliability. In this study, we will thus include agriculture, manufacturing and services in the term \textit{industry} and \textit{industry sector}. In this study we do not consider transportation, neither as sector, nor as end-use within the industry sector.

\cite{VM2015} have analysed the energy use in Danish industry in a high level of detail. They provide the yearly energy demand of 57 different sectors for 20 types of fuel and 22 types of end-use for the year 2012. The distribution of fuels amongst the process categories is also provided. As refinery, public service and construction are not included in the list of sectors for which detailed end-use data is provided \cite{VM2015}, they will not be included in this study.

%The 22 energy end-uses can be clustered in four groups: \textit{internal conversion losses}, \textit{process heating} (heating/boiling, drying, inspissation, distillation, burning/sintering, melting/casting, other process heating up to/over 150 C), \textit{space heating} and \textit{utilities} (energy for heat pumps, lightning, pumping, comfort cooling, cooling/freezing, comfort ventilation, blowers, compressed air, hydraulic machinery, other electric motors, IT and electronics, other electric consumption).
%\\
%Twenty fuel types supplying processes include coal, fuel oil, petroleum coke, coke, diesel, gasoline (unleaded and coloured), LPG, natural gas, gas-oil, bio-gas, biomass (wood chips, wood pellets and straw), bio-oil, waste, electricity, district heating, solar heating and heat pumps.
%\\
The 57 industrial sectors are clustered in five groups according to similarities in temporal patterns of energy consumption: agriculture, production single shift, production double shift, production triple shift and services. The underlying assumptions behind the classification are based on \cite{VM2016,Wiese2017}.

@@

These categories will be used in the following for illustration. Their practicability with regard to time-related data like demand profiles will further be described in \autoref{sec:profiles}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Energy use in Danish industry}
\label{sec:energy_use}

@@

The fuel consumption among the sectors is distributed unevenly, with electricity (36\%), natural gas (21\%), district heating (15\%), gasoil/kerosene (6.4\%) and petroleum coke (4\%) covering the majority of the consumption, while the other fuels contribute to a minor extent.

\iffalse
\autoref{heatmapagri}, \autoref{heatmapprod}, \autoref{heatmapserv} provide an indication on the fuel-use for the different industrial processes. For the production group the data are reported as a single aggregated cluster.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_agri.png}
\caption{Breakdown of energy consumption by end-use and fuel in agriculture \cite{VM2015}}
\label{heatmapagri} 
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_prod.png}
\caption{Breakdown of energy consumption by end-use and fuel in production \cite{VM2015}}
\label{heatmapprod} 
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{Img/dan_ind/heatmap_serv.png}
\caption{Breakdown of energy consumption by end-use and fuel in services \cite{VM2015}}
\label{heatmapserv} 
\end{figure}
\fi

The heat maps, presented in three clustered groups, provide a useful indication on the most energy-intensive processes and on the most employed fuels. 
In agriculture, most of the fuel consumption is related to heating up to 150\textdegree (51\%); in production, drying (15\%), heating/boiling (16\%) and space heating (11\%) are predominant; while in service, space heating (50\%) and lightning (17\%) together consume more than half of the total fuels used.

%top up with data
In production, there is a high diversity of end-use; the role of natural gas as a fuel is predominant. 
Space heating and electricity dominate the total energy use in the service sector. District heat is the main fuel supplying space heat, followed by natural gas.

Furthermore, the data show an uneven distribution of fuel consumption in the various groups, with electricity (34\%) and gasoil/kerosene (25\%) being predominant in agriculture, electricity (31\%) and natural gas (31\%) in production, and electricity (50\%) and district heating (37\%) dominating the service sector.

This description of the Danish industry constitutes the bases to model this sector and to investigate its decarbonisation options. With a clear overview of the end-use processes and the related fuel consumption, the industry sector can be conveniently integrated in an energy system model. 
The following section proposes a mathematical model for the industry sector in an integrated energy system model.

@@

\section{Methods} \label{meths}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Balmorel: energy system model}

The application of energy system models facilitates the assessment of the industry's role in achieving pathways to decarbonised energy systems. As commodities like electricity and heat are highly related, in both power and industry sector (e.g. due to the use of heat pumps), only a simultaneous consideration can reflect cost-minimal solutions from a system perspective. This allows to investigate what is defined as "socio-economic welfare maximisation" under different conditions. 
Among the tools available in the literature \cite{Connolly2010}, the energy system model Balmorel is adopted for the analysis \cite{balmorel}.
Balmorel is an open source, mostly linear energy system model that optimises investments and operation of power plants, storage devices and transmission lines for geographical area that can be defined by the user \cite{Ravn2001,Wiese2018}. 
The model considers a set of neighbouring countries operating in an interconnected electricity market. Each country is composed by one or several regions. Electricity can be traded and transmitted between connected regions, with limits imposed by given transmission capacity. 
To consider the heat sector, each electricity region is divided into several district heating areas (DH). Heat transmission among such areas is not allowed.  
\\
Balmorel considers time according to years, seasons (often used as weeks) and individual time units (often used as hours). The user can choose the time horizon and time resolution of the analysis depending on the requirements for the specific investigation.
\\
The model allows to simulate scenarios where demand and supply of electricity and heat are balanced. Operation and investments are optimised, considering local generation vs. import/export, price elasticity of the demand and other characteristics typical for energy systems \cite{Munster2012}. 
\\
The model relies on a set of exogenous input data, including existing capacities of electricity and heat generation technologies, transmissions lines and heat and power demand. 
Energy generating technologies include CHP (both back-pressure and extraction), heat pumps, storage devices (for electricity and heat), and renewable based production technologies (hydro, wind, solar).
Additional key assumptions relate to fuel prices and CO2 costs. Furthermore, taxes and support schemes are required for specific analysis on impacts of different policies.
The data for the model can be chosen for personal application in specific analyses.
\\
The model has been previously applied for a wide range of studies, such as integration of renewable technologies in the energy mix \cite{Ball2007}, analysis of market conditions \cite{Jensen2008}, policies implementation \cite{Karlsson2008}, future role of district heating \cite{Munster2012,Munster2010} and impact of energy efficiency technologies in the energy system  \cite{Baldini2016a}.

The main Balmorel model is supplemented with several addons that allow for specific investigations \cite{Wiese2018}. As a comprehensive representation of the industry sector is not included yet, we propose to enhance the model with details about industrial energy use and different heat levels and their respective replacement options.

@@

The energy system model Balmorel considers the industry sector, within the optimisation framework, only as a part of the total heat and electricity consumption. Thus the end-use process specification, previously described, are currently not represented. That means, all fuel and process heat demand not covered by district heat are not accounted for. The black dotted line in \autoref{ensyst} represents the boundaries in the current version of Balmorel, where electricity and district heat demand from industry is considered simply as part of the total energy demand.

@@

The block structure presented in \autoref{ensyst} shows the schematic of the novel structure proposed, to consider the diverse characteristics of the industry sector in the energy system model. The same structure can be applied for each industry group considered.

The additions introduced concern mainly the commodities part of the energy demand: electricity, fuels and heat.\\
% Electricity 
The electricity supply has to satisfy, in every single sector, the demand of the respective end-uses. This accurate representation facilitates the investigation of end-use specific decarbonisation options (e.g. savings) at a later stage, as the new options can be applied end-use specific.

Also, electricity can be either be supplied by the grid, or it can be self-produced in on-site industrial plants. It can also be used to run heat pumps and satisfy the heat demand, or "elevate" the quality of heat, for peculiar processes. \\
% Heat
Heat, previously considered only as district heat demand, now includes three temperature levels: space heating, process heat low temperature and process heat high temperature. For this case study, \autoref{tab:temperature_levels} summarises the characteristics of the heat groups, showing how processes are classified according to temperature levels.
Heat can be upgraded (i.e. the temperature is increased) or downgraded (i.e. the temperature is decreased/adapted) using proper heat pumps. 
This means that one can use:
\begin{itemize}
    \item waste heat from a low temperature process to provide heat to a high temperature process
    \item waste heat from high temperature process directly or through a heat pump in a low temperature process
    \item waste heat from a high temperature process to preheat (partly heat) the high temperature process (depending on the temperature considered for low and high temperature processes). 
    \item heat from space heating to provide heat to a low temperature process
\end{itemize}
The excess heat can be used to pre-heat air/water for other processes (e.g. self production) or can be also injected in a DH network where available (i.e. industrial process in proximity to a DH grid). \\
% Fuels
To better represent reality, fuels (used both as energy or fuel-feedstock purposes) are represented directly in the industrial structure with related conversion technologies. This particular facilitates the investigation of CO2 reduction options (like e.g. fuel replacement or electrification) for the industrial sector. \\
%
The modelling of the industry sector requires adaption or extension of different constraints of the energy system model according to the particular considered. The mathematical formulation presented should thus be considered as an extension of the base mathematical model developed for Balmorel (See manual \cite{balmorel}).
